### 2013-12-13 manuscripts/reply/reply-figures.tex
Incorporating new figures into reply ms.
Incorporating updated figures (the same results presented in a
more concise manner). Also, some edits throughout the text.

\mFigure{../images/coin_flip.pdf}{
    A plot of three beta probability density functions that represent a prior
    (black; $beta(10, 10)$), true posterior (blue; $beta(13, 17)$), and
    empirical Bayes density (red; $beta(16, 24)$) for a dataset of 10 Bernoulli
    trials, three of which are successes.
}{figCoinFlip}

%% Exclusion simulation results
\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
    The propensity of the model-averaging approach of \citet{Hickerson2013} to
    exclude the truth.
    The plots illustrate the number of true \divt{} parameters excluded from analyses of
    simulated datasets where \divt{} for 22 pairs of populations is drawn from an
    exponential distribution, $\divt{} \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which at least one true
    parameter value is excluded from the model preferred by a Bayes factor
    ($p(\divt{} \notin \hat{M})$) is also given.
}{figExclusionSimTau}

\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
    The propensity of the model-averaging approach of \citet{Hickerson2013} to
    exclude the truth.
    The plots illustrate the estimated probability of excluding at least one
    true \divt{} value from analyses of simulated datasets where \divt{} for 22
    pairs of populations is drawn from an exponential distribution, $\divt{}
    \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which there is strong support
    for at least one true parameter value being excluded from the model
    ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is also given.
}{figExclusionSimProb}

@@

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_psi_mode.pdf}{
    The bias of the model-averaging approach of \citet{Hickerson2013} to infer
    clustered divergences in our simulation-based power analyses.
    The plots illustrate the estimated number of divergence events ($\hat{\numt{}}$)
    from analyses of simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
    uniform distributions, $\divt{} \sim U(0, \divt{max})$.
    The estimated probability of the method inferring one divergence event, $p(\hat{\numt{}} = 1)$,
    is given for each \divt{max}.
    Each plot represents 500 simulation replicates analyzed using $5\e6$
    samples from the prior.
}{figPowerPsiMode}

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_omega_prob.pdf}{
    The bias of the model-averaging approach of \citet{Hickerson2013} to
    support one divergence event in our simulation-based power analyses.
    The plots illustrate histograms of the estimated posterior probability that
    the dispersion index of divergence times is less than 0.01 ($p(\vmratio{} <
    0.01 | \ssSpace)$) from analyses of simulated datasets where \divt{} for 22
    pairs of populations is drawn from a series of uniform distributions,
    $\divt{} \sim U(0, \divt{max})$.
    The proportion of simulation replicates that strongly support one
    divergence event, $p(BF_{\vmratio{} < 0.01, \vmratio{} \geq 0.01} > 10)$,
    is given for each \divt{max}.
    Each plot represents 500 simulation replicates analyzed using $5\e6$
    samples from the prior.
}{figPowerOmegaProb}

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_num_excluded.pdf}{
    The propensity of the model-averaging approach of \citet{Hickerson2013} to exclude
    the truth in our simulation-based power analyses.
    The plots illustrate the number of true \divt{} parameters excluded from analyses of
    simulated datasets where \divt{} for 22 pairs of populations is drawn from a series of
    uniform distributions, $\divt{} \sim U(0, \divt{max})$. The proportion
    of simulation replicates in which at least one true parameter value is excluded from
    the model preferred by a Bayes factor ($p(\divt{} \notin \hat{M})$) is
    given for each \divt{max}.
    Each plot represents 500 simulation replicates analyzed using $5\e6$
    samples from the prior.
}{figPowerNumExcluded}

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_prob_exclusion.pdf}{
    The propensity of the model-averaging approach of \citet{Hickerson2013} to
    exclude the truth in our simulatoin-based power analyses.  The plots
    illustrate the estimated posterior probability of excluding at least one
    true \divt{} value from analyses of simulated datasets where \divt{} for 22
    pairs of populations is drawn from a series of uniform distributions,
    $\divt{} \sim U(0, \divt{max})$. The proportion of simulation replicates in
    which there is strong support for at least one true parameter value being
    excluded from the model ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is
    given for each \divt{max}.
    Each plot represents 500 simulation replicates analyzed using $5\e6$
    samples from the prior.
}{figPowerProbExclusion}

\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
}{figSamplingError}

@@

% \mFigure{../../response-redux/results/hickerson/pymsbayes-results/pymsbayes-output/d1/m12345678-combined/mean_by_dispersion.pdf}{
%     Results of reanalysis.
% }{figJointPosterior}

>>>>

\mFigure{../images/coin_flip.pdf}{
    A plot of three beta probability density functions that represent a prior
    (black; $beta(10, 10)$), posterior (blue; $beta(13, 17)$), and empirical
    Bayes density (red; $beta(16, 24)$) for a dataset of 10 coin flips, three
    of which are successes.
}{figCoinFlip}

%% Exclusion simulation results
\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/num_tau_excluded.pdf}{
    Histograms of the number of true divergence times excluded from the model
    preferred by the empirically informed model-averaging approach of
    \citet{Hickerson2013} when applied to simulated datasets in which divergence
    times for 22 pairs of populations is drawn from an exponential
    distribution, $\divt{} \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which at least one true
    parameter value is excluded from the preferred model ($p(\divt{} \notin
    \hat{M})$) is also given.
}{figExclusionSimTau}

\mFigure{../../model-choice/results/m1-1-sim/pymsbayes-results/prob_of_exclusion.pdf}{
    Histograms of the support (estimated posterior probabilities) for excluding
    at least one true divergence time when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simulated
    datasets in which divergence times for 22 pairs of populations is drawn
    from an exponential distribution, $\divt{} \sim Exp(2)$.
    The plots represent (A) unadjusted and (B) GLM-adjusted estimates from 1000
    simulation replicates analyzed using $5\e6$ samples from the prior.
    The proportion of simulation replicates in which there is strong support
    for at least one true parameter value being excluded from the model
    ($p(BF_{\divt{} \notin M, \divt{} \in M} > 10)$) is also given.
}{figExclusionSimProb}

@@

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-4.pdf}{
    The tendency of the empirically informed model-averaging approach of
    \citet{Hickerson2013} to (A--D) infer clustered divergences and (E--H)
    support the extreme model of one divergence when applied to simulated
    datasets in which the divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
    Four of the six \divt{max} we simulated are provided; please see
    Figure~S\ref{figPower6} for a summary of all of the results.
}{figPower4}

\mFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-4.pdf}{
    Histograms of the (A--D) number of true divergence-time parameters excluded
    from the preferred model and the (E--H) posterior probability of excluding
    at least one divergence-time parameter when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simlated
    datasets in which divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
    Four of the six \divt{max} we simulated are provided; please see
    Figure~S\ref{figPowerExclusion6} for a summary of all of the results.
}{figPowerExclusion4}

%% Sampling error
\mFigure{../../response-redux/results/sampling-error/pymsbayes-results/omega_over_sampling.pdf}{
    Traces of the estimated lower and upper limits of the 95\% highest posterior density (HPD)
    interval of \vmratio{} (the dispersion index of divergence times) as 100 million prior
    samples are accumulated. Each pair of points is based on 1000 posterior samples retained
    from the prior. Both (A) unadjusted and (b) GLM-regression-adjusted estimates are shown.
    Prior settings were $\divt{} \sim U(0,10)$, $\meanDescendantTheta{} \sim U(0.0005, 0.04)$,
    and $\ancestralTheta{} \sim U(0.0005, 0.02)$.
}{figSamplingError}

@@

### 2013-12-13 manuscripts/reply/reply-si-figures.tex
Incorporating new figures into reply ms.
Incorporating updated figures (the same results presented in a
more concise manner). Also, some edits throughout the text.

\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_median.pdf}{
    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
    estimate the dispersion index of divergence times (\vmratio{}) from
    analyses of simulated datasets where \divt{} for 22 pairs of populations is
    drawn from a series of uniform distributions, $\divt{} \sim U(0,
    \divt{max})$.
    The proportion of estimates less than the true value of,
    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
    Each plot represents unadjusted median estimates from 500 simulation
    replicates analyzed using $5\e6$ samples from the prior.
}{figPowerAccOmegaMedian}

\siFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power_accuracy_omega_mode_glm.pdf}{
    The accuracy of the model-averaging approach of \citet{Hickerson2013} to
    estimate the dispersion index of divergence times (\vmratio{}) from
    analyses of simulated datasets where \divt{} for 22 pairs of populations is
    drawn from a series of uniform distributions, $\divt{} \sim U(0,
    \divt{max})$.
    The proportion of estimates less than the true value of,
    $p(\hat{\vmratio{}} < \vmratio{})$, is given for each \divt{max}.
    Each plot represents GLM-regression-adjusted mode estimates from 500
    simulation replicates analyzed using $5\e6$ samples from the prior.
}{figPowerAccOmegaModeGLM}

%% Validation results
\siFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
    An assessment of the approximate Bayesian model-averageing approach of
    \citet{Hickerson2013} under the ideal conditions when the prior model
    is correct (i.e., the pseudo-replicate datasets are simulated from
    parameters drawn from the same prior distributions used in the
    analysis).
    The plots show the relationship between the estimated posterior and true
    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
    on 50,000 simulations.
    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
    posterior estimate from each simulation replicate.
    The prior settings for all replicates included five prior models with
    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
    U(0.0001, 0.05)$ for all five models, and
    $M_1: \divt{} \sim U(0, 0.1)$,
    $M_2: \divt{} \sim U(0, 1)$,
    $M_3: \divt{} \sim U(0, 5)$,
    $M_4: \divt{} \sim U(0, 10)$, and
    $M_5: \divt{} \sim U(0, 20)$.
    The number of samples from the prior was $2.5\e6$.
    The simulated data structure was 8 population pairs, with a single 1000
    bp locus sampled from 10 individuals from each population.  The 50,000
    estimates of the posterior probability of one divergence event were
    assigned to 20 bins of width 0.05.
    The estimated posterior probability of each bin is plotted against the
    proportion of replicates in that bin with a true value consistent with
    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
}{figValidationMCBehavior}

>>>>

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-accuracy-6.pdf}{
    The accuracy of (A--F) unadjusted and (G--L) GLM-adjusted estimates of
    dispersion index of divergence times (\vmratio{}) when the empirically
    informed model-averaging approach of \citet{Hickerson2013} is applied to
    simlated datasets in which divergence times are randomly drawn from the
    uniform distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top
    of each column of plots (divergence-time distributions are given in units
    of millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
}{figPowerAccuracy}

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-6.pdf}{
    The tendency of the empirically informed model-averaging approach of
    \citet{Hickerson2013} to (A--F) infer clustered divergences and (G--L)
    support the extreme model of one divergence when applied to simulated
    datasets in which the divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
}{figPower6}

\siSidewaysFigure{../../model-choice/results/power-1/pymsbayes-results/plots/power-exclusion-6.pdf}{
    Histograms of the (A--F) number of true divergence-time parameters excluded
    from the preferred model and the (G--L) posterior probability of excluding
    at least one divergence-time parameter when the empirically informed
    model-averaging approach of \citet{Hickerson2013} is applied to simlated
    datasets in which divergence times are randomly drawn from the uniform
    distributions $\divt{} \sim U(0, \divt{max})$ indicated at the top of each
    column of plots (divergence-time distributions are given in units of
    millions of generations ago (MGA) assuming a per-site rate of 1\e{-8}
    mutations per generation).
}{figPowerExclusion6}

%% Validation results
\siFigure{../../model-choice/results/validation/results/pymsbayes-results/plots/mc_behavior.pdf}{
    An assessment of the approximate Bayesian model-averageing approach of
    \citet{Hickerson2013} under the ideal conditions when the prior model
    is correct (i.e., the datasets are simulated from parameters drawn from the
    same prior distributions used in the analysis).
    The plots show the relationship between the estimated posterior and true
    probability of (A \& C) $\numt{}=1$ and (B \& D) $\vmratio{} < 0.01$, based
    on 50,000 simulations.
    The results summarize the (A \& B) unadjusted and (C \& D) GLM-adjusted
    posterior estimate from each simulation replicate.
    The prior settings for all replicates included five prior models with
    $\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
    U(0.0001, 0.05)$ for all five models, and
    $M_1: \divt{} \sim U(0, 0.1)$,
    $M_2: \divt{} \sim U(0, 1)$,
    $M_3: \divt{} \sim U(0, 5)$,
    $M_4: \divt{} \sim U(0, 10)$, and
    $M_5: \divt{} \sim U(0, 20)$.
    The number of samples from the prior was $2.5\e6$.
    The simulated data structure was 8 population pairs, with a single 1000
    bp locus sampled from 10 individuals from each population.  The 50,000
    estimates of the posterior probability of one divergence event were
    assigned to 20 bins of width 0.05.
    The estimated posterior probability of each bin is plotted against the
    proportion of replicates in that bin with a true value consistent with
    one divergence event (i.e., $\numt{}=1$ or $\vmratio{} < 0.01)$.
}{figValidationMCBehavior}

### 2013-12-13 manuscripts/reply/short-reply.tex
Incorporating new figures into reply ms.
Incorporating updated figures (the same results presented in a
more concise manner). Also, some edits throughout the text.

We find that the approach of \citet{Hickerson2013} struggles to estimate the
variance of divergence times (\vmratio{}) across most of the \divt{max}
we simulated, whether evaluating the unadjusted
(Figure~S\ref{figPowerAccOmegaMedian}) or GLM-adjusted
(Figure~S\ref{figPowerAccOmegaModeGLM}) posterior estimates.
The method only estimates \vmratio{} relatively well when the simulated
distribution of divergence times is identical to one of the prior models
(Figures~S\ref{figPowerAccOmegaMedian}E \& S\ref{figPowerAccOmegaModeGLM}E).
This is consistent with the findings of \citet{Oaks2012}  and
\citet{Hickerson2013} that \msb is highly sensitive to the prior distribution
deviating from the true, underlying distribution of the data.

Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
consistently infers highly clustered divergences across all the \divt{max} we
simulated (Figure~\ref{figPowerNumExcluded}).
The method is most likely to infer the extreme case of a single divergence
event shared across 22 taxa when populations diverged randomly over the past
\globalcoalunit generations.
When divergences are random over the past $8\globalpopsize$ generations, the
most likely inference is only two divergence events, and a single
divergence is still estimated in more than 10\% of the replicates.
It is interesting to note that as \divt{max} increases, but before the
estimates are finally pulled away from $\numt{} = 1$, the distribution of
\numt{} estimates closely mirror the U-shaped prior on divergence models
used by \msb (see Figure~\ref{figPowerNumExcluded}E and Oaks et al.'s
(\citeyear{Oaks2012}) Figure 5B).
This suggests this U-shaped prior coupled with the small marginal likelihoods
of models with many divergence parameters, is a major cause of the method's
bias toward clustered divergence models; this is consistent with
\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.

Looking at our simulation results in terms of the posterior probability of the
dispersion index of divergence times supporting the extreme case of one
divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
method strongly supports one divergence in greater than 27\% of the replicates
across all the \divt{max} we simulated (Figure~\ref{figPowerOmegaProb});
following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
the criterion for incorrect inference of a single divergence event.
Furthermore, there is strong support for a single divergence event in
more than 90\% of the replicates when divergences are random over the past
$2.4\globalpopsize$ generations, and more than 60\% when over the past
$3.2\globalpopsize$ generations (Figure~\ref{figPowerOmegaProb}).

@@

Also, the results of our power analyses further demonstrate the propensity of
Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
parameter values.
Across all but one of the \divt{max} we simulated, in a large proportion of
replicates the method favors a model that excludes the truth, and across many
of the \divt{max} the preferred model will exclude a large proportion of the
true divergence-time values (Figure~\ref{figPowerNumExcluded}).
Importantly, the posterior probability of excluding at least one true
divergence value is also quite high across many of the \divt{max}
(Figure~\ref{figPowerProbExclusion}).
Only when the data are identically distributed as one of the prior models does
the method avoid excluding the truth more than 5\% of the time
(Figure~\ref{figPowerNumExcluded}E).
Again, this demonstrates the method's sensitivity to priors.

@@

This is further demonstrated by the results presented herein that show the
model-averaging approach of \citet{Hickerson2013} prefers models with narrower
\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerNumExcluded,figPowerProbExclusion})
and models with fewer \divt{} parameters (Figs.~\ref{figPowerPsiMode} \&
\ref{figPowerOmegaProb}).
In all of these analyses, each of the prior models have the same number of
samples.
Thus, while sampling error will always be present in any numerical Bayesian
approximation method, insufficient sampling is an unlikely explanation for
support of prior models with less parameter space.
While analyses that sample each model proportional to their parameter space
could be explored, it is clear that the marginal likelihoods under broad
uniform priors on divergence times will be greater for models with fewer
divergence-time dimensions.

\section*{General thoughts on the model of \msb}
Our findings are not very surprising in light of difficult inference problem
\msb seeks to tackle.
To get a sense of the challenge \msb is faced with, we tally up all the free
parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
% While this approach is trivial to implement, it is not a trivial change
% to the \msb model.
% To better understand the reason for this, it might help to step back and get a
% sense of the scale of the \msb model.
% To do this, we will use the dataset of 22 vertebrate taxon pairs of
% \citet{Oaks2012} as an example.
% Following the model description and notation of \citet{Oaks2012}, let us tally
% up all of the free parameters in the \msb model.
Under the simplest model in \msb (i.e., assuming no migration and no
intra-locus recombination), the number of parameters for each taxon pair
include:
The population sizes of the ancestral and descendant populations
(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
the magnitude of population contraction in each of the descendant
populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
timing of these contractions (\bottleTime{}), and the $N-1$ node heights
(coalescent times) of the gene tree that gave rise to the $N$ gene
copies sampled from both populations.
% Because we only have a single locus for each taxon, there are no locus-specific
% \myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
% \locusRateHetShapeParameter shape parameter for the gamma prior distribution on
% \locusMutationRateScalar{}.
Lastly, there are between one and 22 divergence time parameters \divt{} in
the vector \divtvector.
Overall, when applying the simplest model in \msb to the dataset of
\citet{Oaks2012} with 22 taxon pairs, there are 581--602 free parameters
(depending on the number of divergence-time parameters in \divtvector).
Furthermore, under this rich model, the method is estimating the probability of
1002 divergence models \citep[i.e., the number of integer partitions of
$Y=22$;][]{Oaks2012}.

This is a very difficult inference problem, especially considering that all the
information in the sequence alignment of each taxon pair is distilled into four
summary statistics:
$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
That gives us a total of 88 summary statistics (four from each of the 22 taxon
pairs), which contain minimal information about many of the $\approx 600$
parameters in the model.
More summary statistics can be used in \msb, but most are highly correlated
with the four default statistics, and thus contribute little additional
information about the parameters from the sequence data.
The large number of parameters and divergence models relative to summary
statistics is undoubtedly a key reason the method lacks robustness to prior
conditions.
Robustness is an important characteristic of a method to gauge its
applicability to real-world data, because we know the model and priors will be
wrong to some degree.

The model-averaging approach of \citet{Hickerson2013} adds an additional
dimension of model choice to the model.
They expand the model to sample over eight prior models.
This adds an additional free parameter to the model and, more importantly,
forces the model to sample over 8016 unique models.
While this approach is trivial to implement, it is a non-trivial extension of
the model.
% Given the method's very large number of free parameters and models to sample
% relative to the information used from the data, is likely a major reason the
% method is so sensitive to priors, and tends to support models with the least
% parameter space when used with uniform priors.
In theory, the model-averaging approach of \citet{Hickerson2013} is very
appealing.
It leverages a great strength of Bayesian statistical procedures, namely the
ability to obtain marginalized estimates that incorporate uncertainty in
nuisance parameters.
However, when averaging over both narrow-empirical and diffuse uniform priors
for a parameter that is expected to have a very non-uniform likelihood density,
and in the context of a model-choice method that is highly sensitive to priors,
it is not too surprising to find the approach struggles.

>>>>

We find that the approach of \citet{Hickerson2013} struggles to estimate the
variance of divergence times (\vmratio{}) across most of the \divt{max}
we simulated, whether evaluating the unadjusted
(Figure~S\ref{figPowerAccuracy}A--F) or GLM-adjusted
(Figure~S\ref{figPowerAccuracy}G--L) posterior estimates.
The method only estimates \vmratio{} relatively well when the simulated
distribution of divergence times is identical to one of the prior models
(Figures~S\ref{figPowerAccuracy}E\&K).
This is consistent with the findings of \citet{Oaks2012}  and
\citet{Hickerson2013} that \msb is highly sensitive to the prior distribution
deviating from the true, underlying distribution of the data.

Furthermore, our results demonstrate that the approach of \citet{Hickerson2013}
consistently infers highly clustered divergences across all the \divt{max} we
simulated (Figure~\ref{figPower4}A--D \& S\ref{figPower6}A--F).
The method is most likely to infer the extreme case of a single divergence
event shared across 22 taxa when populations diverged randomly over the past
\globalcoalunit generations.
When divergences are random over the past $8\globalpopsize$ generations, the
most likely inference is only two divergence events, and a single
divergence is still estimated in more than 10\% of the replicates.
It is interesting to note that as \divt{max} increases, but before the
estimates are finally pulled away from $\numt{} = 1$, the distribution of
\numt{} estimates closely mirror the U-shaped prior on divergence models
used by \msb (see Figure~\ref{figPower4}E and Oaks et al.'s
(\citeyear{Oaks2012}) Figure 5B).
This suggests this U-shaped prior coupled with the small marginal likelihoods
of models with many divergence parameters, is a major cause of the method's
bias toward clustered divergence models; this is consistent with
\citet{Oaks2012} and confirmed by \citet{Oaks2014dpp}.

Looking at our simulation results in terms of the posterior probability of the
dispersion index of divergence times supporting the extreme case of one
divergence event (i.e., $p(\vmratio{} < 0.01 \given \ssSpace)$), we find the
method strongly supports one divergence in greater than 27\% of the replicates
across all the \divt{max} we simulated (Figure~\ref{figPower4}E--H \&
S\ref{figPower6}G--L);
following \citet{Hickerson2013}, we use a Bayes factor of greater than 10 as
the criterion for incorrect inference of a single divergence event.
Furthermore, there is strong support for a single divergence event in
more than 90\% of the replicates when divergences are random over the past
$2.4\globalpopsize$ generations, and more than 60\% when over the past
$3.2\globalpopsize$ generations (Figure~\ref{figPower4}E--H \&
S\ref{figPower6}G--L).

@@

Also, the results of our power analyses further demonstrate the propensity of
Hickerson et al.'s (\citeyear{Hickerson2013}) approach to exclude true
parameter values.
Across all but one of the \divt{max} we simulated, in a large proportion of
replicates the method favors a model that excludes the truth, and across many
of the \divt{max} the preferred model will exclude a large proportion of the
true divergence-time values (Figure~\ref{figPowerExclusion4}A--D \&
S\ref{figPowerExclusion6}A--F).
Importantly, the posterior probability of excluding at least one true
divergence value is also quite high across many of the \divt{max}
(Figure~\ref{figPowerExclusion4}E--H \& S\ref{figPowerExclusion6}G--L).
Only when the data are identically distributed as one of the prior models does
the method avoid excluding the truth more than 5\% of the time
(Figure~\ref{figPowerExclusion4}C).
Again, this demonstrates the method's sensitivity to priors.

@@

This is further demonstrated by the results presented herein that show the
model-averaging approach of \citet{Hickerson2013} prefers models with narrower
\divt{} priors (Table~\ref{tabModelChoiceEmpirical} and
Figs.~\labelcref{figExclusionSimTau,figExclusionSimProb,figPowerExclusion4})
and models with fewer \divt{} parameters (Figure~\ref{figPower4}).
In all of these analyses, each of the prior models have the same number of
samples.
Thus, while sampling error will always be present in any numerical Bayesian
approximation method, insufficient sampling is an unlikely explanation for
support of prior models with less parameter space.
While analyses that sample each model proportional to their parameter space
could be explored, it is clear that the marginal likelihoods under broad
uniform priors on divergence times will be greater for models with fewer
divergence-time dimensions.

\section*{General thoughts on the model of \msb}
Our findings are not surprising in light of the difficult inference problem
with which \msb is faced.
To get a sense of the difficulty of this problem, we tally up all the free
parameters of the \msb model as applied to the dataset of \citet{Oaks2012}.
% While this approach is trivial to implement, it is not a trivial change
% to the \msb model.
% To better understand the reason for this, it might help to step back and get a
% sense of the scale of the \msb model.
% To do this, we will use the dataset of 22 vertebrate taxon pairs of
% \citet{Oaks2012} as an example.
% Following the model description and notation of \citet{Oaks2012}, let us tally
% up all of the free parameters in the \msb model.
Under the simplest model in \msb (i.e., assuming no migration and no
intra-locus recombination), the number of parameters for each taxon pair
include:
The population sizes of the ancestral and descendant populations
(\ancestralTheta{}, \descendantTheta{1}{}, \descendantTheta{2}{}),
the magnitude of population contraction in each of the descendant
populations (\bottleScalar{1}{} and \bottleScalar{2}{}), the
timing of these contractions (\bottleTime{}), and the $N-1$ node heights
(coalescent times) of the gene tree that gave rise to the $N$ gene
copies sampled from both populations.
% Because we only have a single locus for each taxon, there are no locus-specific
% \myTheta{}-scaling parameters (\locusMutationRateScalar{}) or
% \locusRateHetShapeParameter shape parameter for the gamma prior distribution on
% \locusMutationRateScalar{}.
Lastly, there are between one and 22 divergence time parameters \divt{} in
the vector \divtvector.
Overall, when applying the simplest model in \msb to the dataset of
\citet{Oaks2012} with 22 taxon pairs, there are 581--602 free parameters
(depending on the number of divergence-time parameters in \divtvector).
Furthermore, under this rich model, the method is estimating the probability of
1002 divergence models \citep[i.e., the number of integer partitions of
$Y=22$;][]{Oaks2012}.

This is a very difficult inference problem, especially considering that all the
information in the sequence alignment of each taxon pair is distilled into four
summary statistics:
$\pi$ \citep{Tajima1983}, $\theta_W$ \citep{Watterson1975}, $\pi_{net}$
\citep{Takahata1985}, and $SD(\pi-\theta_W)$ \citep{Tajima1989}.
That gives us a total of 88 summary statistics (four from each of the 22 taxon
pairs), which contain minimal information about many of the $\approx 600$
parameters in the model.
More summary statistics can be used in \msb, but most are highly correlated
with the four default statistics, and thus contribute little additional
information about the parameters from the sequence data.
The large number of parameters and divergence models relative to the amount of
information in the data is undoubtedly a key reason the method lacks robustness
to prior conditions.
Robustness is an important characteristic of a method to gauge its
applicability to real-world data, because we know the model and priors will be
wrong to some degree.

The model-averaging approach of \citet{Hickerson2013} adds an additional
dimension of model choice to the model.
They expand the model to sample over eight prior models.
This adds an additional free parameter to the model and, more importantly,
forces the model to sample over 8016 unique models.
While this approach is trivial to implement, it is a non-trivial extension of
the model.
% Given the method's very large number of free parameters and models to sample
% relative to the information used from the data, is likely a major reason the
% method is so sensitive to priors, and tends to support models with the least
% parameter space when used with uniform priors.
In theory, the model-averaging approach of \citet{Hickerson2013} is very
appealing.
It leverages a great strength of Bayesian statistical procedures, namely the
ability to obtain marginalized estimates that incorporate uncertainty in
nuisance parameters.
However, when averaging over both narrow-empirical and diffuse uniform priors
for a parameter that is expected to have a very non-uniform likelihood density,
and in the context of a model-choice method that is highly sensitive to priors,
it is not too surprising that the approach struggles.
